Author Guidelines for DAGM-OAGM Submission

### 2013-09-07 exact_learning/paper.tex
intermediate

\section{Introduction}
% Remove duplication with pystruct intro
As discussed previously, %TODO in the introduction? in pystruct chapter?
 many computer vision algorithms employ conditional random
field models on pixel or superpixel graphs. These graphs, by nature, contain loops,
making exact prediction and loss-augmented prediction in general intractable.
As loss-augmented prediction is a central step in the learning algorithms
discussed in Chapter~\ref{ch:structured_pystruct}
in the inference lead to only approximate learning.

In this chapter, we investigate the necessity and consequences of these
approximations in the one-slack learning algorithms in the context of semantic
image segmentation. We show that despite inference being deemed intractable in
loopy superpixel models, we are able to learn a pairwise conditional random
field model using a structured support vector machine exactly for this task.
We evaluate our approach on the popular MSRC-21 and Pascal VOC datasets. % and nyu?

The contribution of this chapter is threefold:
\begin{itemize}
    \item We analyze the simultaneous use of multiple approximate inference
        methods for learning SSVMs using the cutting plane method, relating
        approximate learning to the exact optimum.
    \item We introduce an efficient caching scheme to accelerate cutting plane
        training.
    \item We demonstrate that using a combination of under-generating and exact
        inference methods, we can learn an SSVM exactly in a practical
        application, even in the presence of loopy graphs.
    % also fusion moves is pretty good?
\end{itemize}

While empirically exact learning yields results comparable to those using
approximate inference alone, certification of optimality allows treating
learning as a black-box, allowing the researcher to focus attention on
designing the model for the application at hand.

@@

Recently, there has been much research in learning structured prediction models
where standard exact inference techniques are not applicable.
The influence of approximate inference on structural support vector machine
learning was first analyzed by  \citet{finley2008training}.
\citet{finley2008training} show convergence results for under-generating and
over-generating inference procedures, meaning methods that find suboptimal, but
feasible solutions, and optimal solutions from a larger (unfeasible) set,
respectively.
\citet{finley2008training} demonstrate that over-generating approaches---in
particular linear programming (LP)---perform best on the considered model.
They also show that learning parameters with the LP relaxation minimizes a
bound on the empirical risk when extending the target domain to the relaxed
solutions.
In contrast, we aim at optimizing the non-relaxed objective directly, minimizing
the original empirical risk. This is an important difference, as relaxed solutions
are usually not acceptable in practice.

As using LP relaxations was considered too costly for typical computer vision
approaches, later work employed graph-cuts~\citep{szummer2008learning} or Loopy
Believe Propagation (LBP)~\citep{lucchi2011spatial}. These works use a single
inference algorithm for during the whole learning process, and can not provide any
bounds on the true objective or the empirical risk. In contrast, we combine
different inference methods that are more appropriate for different stages of
learning.

@@

\section{Learnint Structure Support Vector Machines with Approximate Inference}
%TODO ueberleitung zu approximate inference
% wir machen one-slack und I ist jetzt das und das.
There are two groups of inference procedures, as identified in
\citet{finley2008training}: under-generating and over-generating approaches.
An under-generating approach satisfies $I(x^i, y^i, \theta) \in
\mathcal{Y}$, but does not guarantee maximality in line~\ref{get_cutting_plane}
of Algorithm~\ref{alg_cutting_plane}. An over-generating approach on the other
hand, will solve the loss-augmented inference in line~\ref{get_cutting_plane}
exactly, but for a larger set $\hat{\mathcal{Y}} \supset \mathcal{Y}$, meaning
that possibly $I(x^i, y^i, \theta) \notin \mathcal{Y}$.  We will elaborate on
the properties of under-generating and over-generating inference procedures in
Section~\ref{bounds}. %TODO Filler sentence? wat?

@@

Let $o_{\W}(\theta)$ denote objective \eqref{oneslack} with
given parameters $\theta$ restricted to a working set $\W$, as computed in
line~\ref{restricted} of Algorithm~\ref{alg_cutting_plane} and  let
\[
    o^I(\theta) = C\xi' + \frac{||\theta||}{2}^2
\]
when using inference algorithm $I$, that is $o^I(\theta)$ is the approximation of the primal
objective given by $I$. To simplify exposure, we will drop the dependency on $\theta$.

@@

We can also use these observations to judge the suboptimality of a given parameter $\theta$,
that is see how far the current objective is from the true optimum.
Learning with any under-generating approach, we can use 1. to maintain a lower bound
on the objective. At any point during learning, in particular if no more constraints
can be found, we can then use 2., to also find an upper bound.
This way, we can empirically bound the estimation error, using only approximate
inference.
We will now describe how we can further use 1. to both speed up and improve learning.

@@

%objective given by $I$. To simplify exposure, we will drop the dependency on $\theta$.
In the following, we will use the notation of Section~\ref{bounds} and indicate
the choices of oracle $I$ with $Q$ for a chosen inference algorithm and $C$ for
using cached constraints.

@@

\subsection{Implementation Details}
We implemented the cutting plane Algorithm~\ref{alg_cutting_plane} and our pairwise
model in Python.
Our cutting plane solver for Equation~\eqref{mainequation} uses
\texttt{cvxopt}~\citep{dahl2006cvxopt} for solving the QP in the inner loop. Code for
our structured prediction framework will be released under an open source
license upon acceptance.

We used the SLIC implementation provided by \citet{achanta2012slic} to extract superpixels and
the SIFT implementation in the \texttt{vlfeat} package~\citep{vedaldi08vlfeat}.
For clustering visual words, piecewise training of unary potentials and the
approximation to the $\chi^2$-kernel, we made use of the \texttt{scikit-learn}
machine learning package~\citep{pedregosa2011scikit}.
The move-making algorithm using QPBO is implemented with the help of the QPBO-I
method provided by \citet{rother2007optimizing}.
We use the excellent implementation of AD$^3$ provided by the authors of
\citet{martins2011augmented}.

>>>>

\section{Introduction}
As discussed previously, %TODO in the introduction? in pystruct chapter?
 many computer vision algorithms employ conditional random
field models on pixel or superpixel graphs. These graphs, by nature, contain loops,
making exact prediction and loss-augmented prediction in general intractable.
As loss-augmented prediction is a central step in the learning algorithms
discussed in Chapter~\ref{ch:structured_pystruct}, approximate inference leads
to complications in learning.

In this chapter, we investigate the necessity and consequences of these
approximations in the one-slack cutting-plane learning algorithm in the context of semantic
image segmentation. We show that despite inference being deemed intractable in
loopy superpixel models, we are able to learn a pairwise conditional random
field model using a structured support vector machine exactly for this task.
We evaluate our approach on the popular MSRC-21 and Pascal VOC datasets. % and nyu?

The contribution of this chapter is threefold:
\begin{itemize}
    \item We analyze the simultaneous use of multiple approximate inference
        methods for learning SSVMs using the cutting plane method, relating
        approximate learning to the exact optimum. %FIXME really?
    \item We introduce an efficient caching scheme to accelerate cutting plane
        training.
    \item We demonstrate that using a combination of under-generating and exact
        inference methods, we can learn an SSVM exactly in a practical
        application, even in the presence of loopy graphs.
    % also fusion moves is pretty good?
\end{itemize}

While empirically exact learning yields results comparable to those using
approximate inference alone, certification of optimality allows treating
learning as a black-box, enabling the researcher to focus attention on
designing the model for the application at hand.

@@

Recently, there has been an increase in research in learning structured
prediction models where standard exact inference techniques are not applicable,
in particular in the computer vision community.
The influence of approximate inference on structural support vector machine
learning was first analyzed by  \citet{finley2008training}.
\citet{finley2008training} show convergence results for under-generating and
over-generating inference procedures, meaning methods that find suboptimal, but
feasible solutions, and optimal solutions from a larger (unfeasible) set,
respectively.
\citet{finley2008training} demonstrate that over-generating approaches---in
particular linear programming (LP)---perform best on the considered model.
They also show that learning parameters with the LP relaxation minimizes a
bound on the empirical risk when extending the target domain to the relaxed
solutions.
In contrast, we aim at optimizing the non-relaxed objective directly, minimizing
the original empirical risk. This is an important difference, as relaxed solutions
are usually not acceptable in practice.

As using LP relaxations was considered too costly for typical computer vision
approaches, later work employed graph-cuts~\citep{szummer2008learning} or Loopy
Believe Propagation (LBP)~\citep{lucchi2011spatial}. These works use a single
inference algorithm for the whole learning process, and can not provide any
bounds on the true objective or the empirical risk. In contrast, in this chapter
we show how to combine different inference methods that are more appropriate
for different stages of learning.

@@

\section{Learning Structure Support Vector Machines with Approximate Inference}
In this chapter, we will use the 1-slack cutting plane algorithm, as we found this
algorithm to be most suitable for out approach. For reference, the algorithm is outlined
\Secref{one_slac}, Algorithm~\ref{alg_one_slack}. The inference algorithm $I$ is called
in line~\ref{get_cutting_plane}. We investigate algorithms $I$ (often called separation
oracles in the context) that will not yield the exact maximum here.
There are two groups of inference procedures, as identified in
\citet{finley2008training}: under-generating and over-generating approaches.
An under-generating approach satisfies $I(x^i, y^i, \theta) \in
\mathcal{Y}$, but does not guarantee maximality in line~\ref{get_cutting_plane}
of Algorithm~\ref{alg_cutting_plane}. An over-generating approach on the other
hand, will solve the loss-augmented prediction in line~\ref{get_cutting_plane}
exactly, but for a larger set $\hat{\mathcal{Y}} \supset \mathcal{Y}$, meaning
that possibly $I(x^i, y^i, \theta) \notin \mathcal{Y}$.
%We will elaborate on
%the properties of under-generating and over-generating inference procedures in
%Section~\ref{bounds}. %TODO Filler sentence? wat?

@@

Let $o_{\W}(\theta)$ denote objective \eqref{oneslack} with
given parameters $\theta$ restricted to a working set $\W$, as computed in
line~\ref{restricted} of Algorithm~\ref{alg_cutting_plane} and  let
\[
    o^I(\theta) = C\xi' + \frac{||\theta||}{2}^2
\]
when using inference algorithm $I$, that is $o^I(\theta)$ is the approximation of the primal
objective given by $I$. To simplify exposure, we drop the dependency on $\theta$.

@@

We can also use these observations to judge the suboptimality of a given parameter $\theta$,
that is see how far the current objective is from the true optimum.
Learning with any under-generating approach, we can use 1. to maintain a lower bound
on the objective. At any point during learning, in particular if no more constraints
can be found, we can then use 2., to also find an upper bound.
This way, we can empirically bound the estimation error, using only approximate
inference.
We now describe how we can further use 1. to both speed up and improve learning.

@@

%objective given by $I$. To simplify exposure, we drop the dependency on $\theta$.
In the following, we use the notation of Section~\ref{bounds} and indicate
the choices of oracle $I$ with $Q$ for a chosen inference algorithm and $C$ for
using cached constraints.

@@

\subsection{Implementation Details}
We implemented the described procedure in \pystruct.

We used the SLIC implementation provided by \citet{achanta2012slic} to extract superpixels and
the SIFT implementation in the \texttt{vlfeat} package~\citep{vedaldi08vlfeat}.
For clustering visual words, piecewise training of unary potentials and the
approximation to the $\chi^2$-kernel, we made use of the \texttt{scikit-learn}
machine learning package~\citep{pedregosa2011scikit}.
%TODO implementation details here?
The move-making algorithm using QPBO is implemented with the help of the QPBO-I
method provided by \citet{rother2007optimizing}.
We use the excellent implementation of AD$^3$ provided by the authors of
\citet{martins2011augmented}.

### 2013-09-07 itm/paper.tex
intermediate

\begin{table}[t]
\centering
\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lrrrccccc}
\toprule
\multicolumn{4}{c}{Dataset} &\multicolumn{4}{c}{Results}\\
\cmidrule{1-4}
\cmidrule{5-9}
Description &     n &    d &   k &  $k$-Means      & MeanNN        & SL            &   ITM        & ITM ID \\
\cmidrule{1-4}                   \cmidrule{5-5}    \cmidrule{6-6}  \cmidrule{7-7} \cmidrule{8-8} \cmidrule{9-9}
digits      &  1797 &   64 &  10 &     0.62 &      0.67            &0.10          &\B{0.85}       &0.73\\ 
faces       &   400 & 4096 &  40 &     0.41 &      0.49            &0.08          &0.02           &\B{0.54}\\ 
iris        &   150 &    4 &   3 &     0.72 &      0.75            &0.55          &\B{0.88}       &0.88\\
usps        &  9298 &  256 &  10 &     0.52 &      0.54            &0.00          &   0.44        &\B{0.64}\\ 
vehicle     &   846 &   18 &   4 &     0.10 &      0.09            &0.00          &\B{0.10}       &0.10\\
vowel       &   990 &   10 &  11 &     0.17 &      0.19            &0.00          &\B{0.20}       &0.19\\
waveform    &  5000 &   21 &   2 & \B{0.37} &      0.30            &0.00          &   0.23        &0.23\\
mnist       & 70000 &  784 &  10 &     0.37 &      N/A$^\dagger$         &0.00          &   0.50        &0.77\\
\bottomrule
\end{tabularx}
\caption{Adjusted Rand Index of $k$-Means, MeanNN, single link agglomerative
    clustering and ITM on several datasets (higher is better). ITM ID refers to
    ITM using the estimated intrinsic dimensionality. The best score for each
    dataset is printed in bold.\\$^\dagger$\footnotesize We were unable to make
    MeanNN scale to 70000 data points, as storing the whole pairwise distance
    matrix seems critical.\label{results}}
\end{table}

>>>>

\begin{table}[t]
\centering
\begin{tabularx}{\linewidth}{@{\extracolsep{\fill}}lrrrccccc}
\toprule
\multicolumn{4}{c}{Dataset} &\multicolumn{4}{c}{Results}\\
\cmidrule{1-4}
\cmidrule{5-9}
Description &     n &    d &   k &  $k$-Means      & MeanNN        & SL            &   ITM        & ITM ID \\
\cmidrule{1-4}                   \cmidrule{5-5}    \cmidrule{6-6}  \cmidrule{7-7} \cmidrule{8-8} \cmidrule{9-9}
digits      &  1797 &   64 &  10 &     0.62 &      0.67            &0.10          &\B{0.85}       &0.73\\ 
faces       &   400 & 4096 &  40 &     0.41 &      0.49            &0.08          &0.02           &\B{0.54}\\ 
iris        &   150 &    4 &   3 &     0.72 &      0.75            &0.55          &\B{0.88}       &0.88\\
usps        &  9298 &  256 &  10 &     0.52 &      0.54            &0.00          &   0.44        &\B{0.64}\\ 
vehicle     &   846 &   18 &   4 &     0.10 &      0.09            &0.00          &\B{0.10}       &0.10\\
vowel       &   990 &   10 &  11 &     0.17 &      0.19            &0.00          &\B{0.20}       &0.19\\
waveform    &  5000 &   21 &   2 & \B{0.37} &      0.30            &0.00          &   0.23        &0.23\\
mnist       & 70000 &  784 &  10 &     0.37 &      N/A$^\dagger$         &0.00          &   0.50        &0.77\\
\bottomrule
\end{tabularx}
\caption{Adjusted Rand Index of $k$-Means, MeanNN, single link agglomerative
    clustering and ITM on several datasets (higher is better). ITM ID refers to
    ITM using the estimated intrinsic dimensionality. The best score for each
    dataset is printed in bold.\\$^\dagger$\footnotesize We were unable to make
    MeanNN scale to 70000 data points, as storing the whole pairwise distance
    matrix seems necessary.\label{results}}
\end{table}
